Embedded Domain Specific Languages inside Functional Languages

### 2014-03-13 paper.tex
Starting the loop

There are many places this design pattern can be used. One example is the specification
of surface textures as functions; it is possible to export these into code executable
on GPUs, simultaneously lifting the abstractions used to write textures, and speeding
up how fast a shallow embedding of the same operations would run.  There is nothing
that is specific about Haskell, or even functional languages here. Indeed, the same
ideas have been used in Java, for a VHDL generator. Haskell, with its powerful abstractions,
allows deep DSLs to be more than just function calls, and almost feel like a straightforward
shallow embedding.

@@

Expressing function calls in terms of constructors,  and
building expression trees is a gimic. THere are entire
systems build on the idea of building structures 
in principaled ways, for example JHDL~\cite{..}, but
we can do much more. With careful construction,
we can capture function defintions,
and other synatically structuruals,
directly from a deep embedding. It is at this
point, the idea of capturing code,
the using the captured code to execute code
on a different target becomes possible.

@@

Here we have a function that acts over our new type
\verb|E|, and returns a new \verb|E|. How can we capture
this function.
The trick is to invent a unique E, and pass it as the argument
to \verb|f|.

@@

This is remarkable. We're run a function with 
a dummy argument (called the protypical argument)
and extracted the body of the function.

About where is used.

@@

Lava programs are written as equations of recursive bindings.
If we attempt to build a deep embedding of Lava directly,
we will get into a infinite cycle. In order to illustrate
the challenge, we will build a deep embedding of Lava,
see where it goes wrong, and fix it using a technique called
observable sharing.

@@

data Signal where
  Register :: a -> Signal a                      -> Signal a
  Mux2     :: Signal Bool -> (Signal a,Signal a) -> Signal a
  Var      :: String                             -> Signal a -- the trick

mux2 :: ...
mux2 = mux2

>>>>

@@

Expressing function calls in terms of constructors,  and
building expression trees is a useful, but by itself is a gimic. 
With careful construction, we can capture function definitions,
and other synaxtically structures,
directly from a deep embedding. It is at this
point, the idea of capturing code,
the using the captured code to execute code
on a different target becomes possible.
Consider a simple function to add one to its argument.

@@

Here we have a function that acts over our new type
\verb|E|, and returns a new \verb|E|. How can we capture
this function?
The trick is to invent a unique E, and pass it as a (dummy) argument
to \verb|f|.

@@

This is remarkable! We're run a function with 
a dummy argument (called the prototypical argument)
and extracted the body of the function.

There are many places this design pattern can be used. One example is the specification
of surface textures as functions; it is possible to export these into code executable
on GPUs, simultaneously lifting the abstractions used to write textures, and speeding
up how fast a shallow embedding of the same operations would run.  There is nothing
that is specific about Haskell, or even functional languages here. Indeed, the same
ideas have been used in Java, for a VHDL generator. Haskell, with its powerful abstractions,
allows deep DSLs to almost feel like a straightforward shallow embedding.

@@

Lava programs are written as equations of recursive bindings.
If we attempt to build a deep embedding of Lava directly,
we will get into a infinite cycle of structures. In order to illustrate
the challenge, we will build a deep embedding of Lava,
see where it goes wrong, and fix it using a technique called
observable sharing.

@@

data Signal where
  Register :: a -> Signal clk a                  -> Signal clk a
  Mux2     :: Signal Bool -> (Signal a,Signal a) -> Signal clk a
  Var      :: String                             -> Signal clk a -- the Var trick

mux2 :: Signal clk Bool -> (Signal a,Signal a) -> Signal a
mux2 c (a,b) = Mux2 c (a,b)
